---
title: "PSTAT 131 HW 2"
author: "Jay Shreedhar"
date: '2022-04-07'
output:
    html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br /><br />
```{r, include=FALSE}
library(ggplot2)
library(corrplot)
library(tidyverse)
library(tidymodels)
library(ggthemes)
library(dplyr)
library(tibble)
library(readr)
```

```{r}
tidymodels_prefer()
abalone <- read.csv("abalone.csv")
set.seed(3945)
```

**Question 1:**

```{r}

abalone <- abalone %>%
  mutate(age = rings + 1.5)

abalone %>% 
  ggplot(aes(x = age)) +
  geom_histogram(bins = 150) +
  theme_bw()

```
<br />
The histogram of *age* shows that the data forms a bell curve skewed to the right, with the majority of the values falling around 10. 
<br /><br />
**Question 2:**
```{r}

abalone_split <- initial_split(abalone, prop=0.8, strata=age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)

```
<br /><br />
**Question 3:**
```{r}

abalone_recipe <- recipe(age ~ ., data=abalone_train %>% select(-rings))  %>%
  step_dummy(all_nominal_predictors())%>% 
  step_interact(terms = ~ starts_with("type"):shucked_weight) %>%
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms = ~ shucked_weight:shell_weight) %>% step_normalize() %>%   step_scale() %>% step_center(all_predictors())

```
<br /><br />
**Question 4:**
```{r}

lm_model <- linear_reg() %>%
  set_engine("lm")

```
<br /><br />
**Question 5:**
```{r}

lm_workflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(abalone_recipe)

```
<br /><br />
**Question 6:**
```{r}

lm_fit <- fit(lm_workflow, abalone_train)
abalone_predict <- data.frame(type = 'F', longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)
predict(lm_fit, new_data = abalone_predict)

```
The model predicts 24.84932.
<br /><br />
**Question 7:**
```{r}

library(yardstick)
abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_tibble <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_tibble <- bind_cols(abalone_tibble, abalone_train %>% select(age))
abalone_metrics(abalone_tibble, truth = age, estimate = .pred)


```
The RMSE is 2.1457732. The standard $R^2$ is 0.5566857. The MAE is 1.5493796. This $R^2$ value means that 55.66% of the variation in age is explained by the other predictors. This isn't a very high value, meaning that the other predictors are not useful in determining hypothetical values of abalone age.	